A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans

### 2023-07-24 manuscript/manuscript.tex
ms updates from italy

    Across the human genome, there are large-scale fluctuations in genetic
    diversity caused by the indirect effects of selection. This can be thought
    of as a ``linked selection signal" that reflects the effects of selection
    varying according to the placement of functional regions and recombination
    rates along the genome. Previous work has shown that negative selection
    against the steady influx of new deleterious mutations into conserved
    regions is the predominant mode of selection in humans. However, the
    theoretic model that underpins these results, classic Background Selection
    theory, is only applicable when new mutations are so deleterious that they
    cannot drift up in frequency and ultimately fix in the population. Here, we
    develop and apply a statistical method to fit the linked selection signal
    based on an alternate theoretic approach. This theory is based on a
    quantitative genetics view of the linked selection process, which models
    the cumulative effects of weak draft created according the arrangement of
    polygenic additive fitness variance along the genome. Using a recent model
    that jointly predicts the equilibrium fitness variance and substitution
    rates due to deleterious mutations, we estimate the distribution of fitness
    effects (DFE) and mutation rate across three human populations. While our
    model can accommodate weaker selection, we initially find evidence across
    three human populations of very strong selection against deleterious
    mutations consistent with previous work. However, the corollary predictions
    of the substitution rates from our model suggest an unreasonably low rate
    of divergence in highly-conserved region which disagree with observations.
    We hypothesize that this is driven by selected sites within conserved
    segments experiencing a further reduced effective population size due to
    selection across other segments. Extending our method to approximate these
    effects through a locally-reduced effective population size brings
    predicted substitution rates into agreement with observed divergence rates
    and revises DFE estimates to include a sizable portion of weakly
    deleterious mutations. However, this also leads to implausibly large
    mutation rate estimates. Overall, while our model of the genomic linked
    selection signal brings us a step towards uniting population and
    quantitative genetic selection models with the substitution process, our
    work suggests considerable uncertainty remains about the processes
generating fitness variance in humans.

@@

Similar to previous work, we aim to characterize selection processes through
the signal left on nucleotide diversity at linked neutral sites. Although
nucleotide diversity is a less informative statistic than the full site
frequency spectrum, it is much easier to derive theoretic models of how linked
selection alters expected pairwise coalescence times than to describe the
entire genealogy. Still, inferences drawn from the linked selection signal
about selective processes are only as accurate as the underlying theoretic
model. Since we base our method on a fundamentally different theoretic approach
to modeling linked selection, we briefly introduce the theory was extend to
apply across whole-genomes.

Our work extends quantitative genetic models of linked selection
\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
which model the effects linked selection in terms of polygenic additive fitness
variance ($V_A$). These linked selection models stem from
\textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
fitness variation generates weak genetic draft. At the individual level
Robertson considered, selection generates autocorrelation as offspring from
large families tend to beget many descendents themselves (and likewise with
small families) when fitness is heritable. This same autocorrelation occurs at
the genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as
the perturbations to a neutral allele's trajectory from its particular fitness
background tend to occur in the same direction across generations until the
background recombines off. These models quantify the total impact of the
autocorrelation generated by weak draft in terms of what we think of as a
\emph{draft-effective} population size $N_d$. The key insight is that in the
long-run, steady, weak draft created by the presence of additive genetic
fitness variance $V_A > 0$ acts like an extra variance in offspring number
would under pure drift \parencite{Wright1938-tv}. However, due to
autocorrelation, the cumulative effect of this fitness variance is inflated by
a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the expected
total variance in reproductive success a neutral mutation experiences over its
lifetime in a system with weak selection draft.

@@

\begin{align}
    \label{eq:polygenic_Ne}
    N_d \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
\end{align}
%
where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Materials Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. Under equilibrium levels of fitness variation, the
autocorrelation function for a neutral allele associated with segment $i$ is
$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
the segment and $\kappa_i$ is the rate that the associated fitness variance
decays due to selective dynamics. Then, the cumulative autocorrelation is,

\begin{align}
    \label{eq:Q}
    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \nonumber \\
        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)}.
\end{align}

(see Supplementary Materials Equation \ref{supp-eq:Qinf}). This general equation can accommodate
models of polygenic selection as long as the equilibrium additive fitness
variation $V_{A,i}$ can be specified and the change in variance due to
selection can be approximated as a geometric decay, i.e. $\Delta V_{A,i} =
-\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This
is usually a reasonable assumption since within-generation selection removes a
fraction of phenotypic variation from the population, and some fraction of that
is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.

@@

\begin{align}
    N_d = N \exp \left( - \sum_i^S \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
\end{align}

which is identical to the genome-wide model of background selection used in
previous studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}.
Thus, the classic background selection model is a special case of the more
general theory of Santiago and Caballero (\citeyear{Santiago2016-mu}), which
they had shown previously (\citeyear{Santiago1998-bs}).

However, when new mutations are only weakly deleterious, they can drift up in
frequency before their eventual loss or fixation. At this point, the number of
deleterious mutations is no longer well-approximated by the deterministic
mutation-selection-balance theory; their dynamics are strongly influenced by
stochastic perturbations due to both drift and draft. In this weak selection
regime, classic BGS no longer accurately predicts levels of linked diversity
\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
Moreover, selection against weakly deleterious mutations alters the
genealogies, such that they are no longer well-approximated by a rescaled
neutral coalescent as assumed under the classic BGS model
\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. To further
complicate matters, the distribution of the number of deleterious mutations
(and its corresponding fitness distribution) is no longer a stationary Poisson
distribution, instead becoming a traveling wave
\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
population fitness (since we ignore beneficial mutations). When this occurs,
individuals in the least-loaded class are stochastically lost, fixing those
mutations with a click of ``Muller's ratchet"
\parencite{Muller1964-ki,Charlesworth1997-qn}. Unfortunately, determining the
rate of Muller's ratchet is another difficult, open problem
\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz,Higgs1995-xc} related to
Hill--Robertson interference \parencite{Felsenstein1974-xm}.

Under qualitative genetic models of linked selection, these complications are
avoided by finding a general expression for the equilibrium fitness variance
that is applicable for both strong and weakly deleterious mutations.
\textcite{Santiago2016-mu} suggest that a general equation for equilibrium
fitness variance must account for the chance fixations of deleterious mutations
that can occur. Each fixation (or click of Muller's ``ratchet") decreases the
fitness variance, as a previously segregating polymorphism fixes and its
heterozygosity becomes zero. If we let $R$ represent the fixation rate (i.e.
ratchet rate) in the region per generation, each fixation removes the
equivalent amount of equilibrium variation put in by mutation. Thus, the
steady-state genic variance under mutation and negative selection is (omitting
the segment index $i$ for clarity),

@@

where the condition $V_a \ge 0$ is met when the probability of fixation is less
than or equal to the neutral fixation probability of $\nicefrac{1}{2N}$, which
is held when considering deleterious mutations. This equation connects the
equilibrium additive genic fitness variance to the flux of new variation in due
to new deleterious mutations and the flux out due to their substitution and the
decline in mean population fitness. When $R=0$, selection is so strong it
cannot fix, and the equilibrium fitness variation is due entirely to young rare
mutations before their extinction $V_a = V_a^{BGS} \approx Us$. Santiago and
Caballero derive Equation \eqref{eq:Va} through Fisher's Fundamental Theorem of
Natural Selection, but we find an alternative proof (\cite{Higgs1995-xc}; see
Supplementary Materials Section \ref{supp:weak-strong}). We also find that the
steady-state additive genic variance in Equation \eqref{eq:Va} results from
diffusion models with a flux of mutations into discrete sites
(\cite{Kimura1969-jw}).

@@

Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
ratchet rate for a single segment under general mutation-selection processes,
we next extended their model so that it could be applied to whole-genome data.
Our software method \texttt{bgspy} numerically solves Equations
\eqref{eq:main_eqns} to compute the equilibrium additive genic fitness variance
($V_a$) and ratchet rate ($R$) across grids of mutation rates and selection
coefficients. This is done for each pre-specified segment in the genome that
may be under negative selection (e.g. coding sequences or UTRs). We call the
set of predicted reductions across these grids the B' maps (to distinguish them
from McVicker's B maps); these can be used then to find the equilibrium
reduction factor $B(x)$ for any genomic position $x$.

We validated our predicted B' reduction maps with realistic chromosome-scale
forward simulations of negative selection in conserved regions. Averaging over
$100$ replicates, we estimated the simulation empirical reduction map and
compared this to the predicted B and B' maps for the same parameters. We find
that our B' maps and the classic BGS theory B maps closely match simulations
when selection is strong (top row of Figure \ref{fig:figure-1}E). This provides
the first realistic forward-time chromosome-scale simulation confirmation of
classic BGS theory. However, we find slight discrepancies in regions with low
recombination (Figure \ref{fig:figure-1}E). Second, we find our theory is
vastly more accurate than the classic B maps when selection is very weak ($2N_e
s \ll 1$; bottom row of Figure \ref{fig:figure-1}E), which is expected since
BGS theory does not work in this domain. Across all mutation and selection
parameters simulated, the relative error of the classic B maps is 14.6\%
whereas the relative error in the new B' maps is 5\%. Additionally, we find
that the mean squared error between simulations and B' maps is close to the
theoretic lower bound set by the coalescence variance
\parencite{Tajima1983-gu}.

The discrepancy between our theoretic B' maps and chromosome-wide simulations
that leads to the 5\% error occurs entirely around the drift-barrier domain of
$2Ns \approx 1$ (Figure \ref{fig:figure-1}D). The higher error in this domain
occurs across all mutation rates. We hypothesized this error may be due to
selective interference between segments that is not taken into account when we
numerically solve Equations \eqref{eq:main_eqns} for each segment
independently. In particular, we use a fixed drift-effective $N=1000$
corresponding to the number of diploids in the simulations rather than $B(x)
N$, which is an approximation to the draft-effective population size
experienced by a selected segment at position $x$. To test this, we implemented
a ``locally-rescaled" version of the B' maps, which numerically solves
Equations \eqref{eq:main_eqns} for each segment at position $x$ taking into
account the reduction experienced due to selection at other segments by setting
$B(x)N$ as the population size. We find the locally-rescaled B' maps reduce the
relative error from 5\% to 0.4\% and mean squared error (Figure
\ref{fig:figure-1}D, dashed colored lines), but does not entirely eliminate the
error in the $2Ns \approx 1$ domain. Overall, while we find evidence that
local-rescaling slightly reduces error in the $2Ns \approx 1$ domain, we do not
use it during in our genome-wide inference method. This is because to avoid
circularity and possible overfitting, our method would need to solve the
equilibria equations for all segments simultaneously; unfortunately, this is
currently computational unfeasible.

Finally, we hypothesized that the remaining error is because of the build up of
negative linkage disequilibria between selected sites due to Hill--Robertson
interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. To assess
this possibility, we calculated the sum of all linkage disequilibria in our
chromosome-wide simulations. We find negative linkage disequilibria build ups
around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is stronger
when mutation rates are higher. As $Ns \to 0$, the variance in LD inflates as
expected \parencite{Ohta1969-ae,Hill1968-ue}. Overall, the build up of negative
LD is consistent our view that the equilibrium fitness variance modeled by the
SC16 theory is the additive \emph{genic} fitness variance, which differs from
the additive genetic variance by the sum of linkage disequilibria between
selected sites (i.e. $\delta_{LD} = s^2 \sum_{i\ne j} D_{i,j}$ where $D_{i,j}$
is the LD between sites $i$ and $j$). However, according to theory, the
reductions in diversity should be determined by levels of additive genetic
fitness variance that include the contribution of LD (Supplementary Materials
Section \ref{supp:theory}).

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=0.8\textwidth]{figures/figure_2.pdf} 
    \caption{Comparison of parameter estimates using classic BGS theory (green
        lines) with our new B' method (blue lines) across both full and sparse
        track types (dark versus light hue), and different mutation rates
        (columns). In the first row, $R^2$ between predictions and observations
        increases with selection intensity. Both classic BGS and B' methods fit
        the data equally well with sparse tracks, though the estimated
        selection coefficients are inaccurate when selection is weak (second
        row). Mutation rate estimates (third row) are more accurately estimated
        by the B' method than classic BGS, but overall show slight biases.
        Overall, classic BGS methods fail as expected when full-coverage tracks
    are used, since it cannot accommodate weak selection and neutrality. }
  \label{fig:figure-2}
\end{figure}

@@

First, we find the $R^2$ between predicted and simulated megabase-scale
diversity serves as a measure of the strength of the linked selection signal in
genome-wide data. The $R^2$ increases with the intensity of selection against
new deleterious mutations and mutation rate (Figure \ref{fig:figure-2}, top
row). Under just drift or weak negative selection, the variance in diversity is
driven by unstructured coalescence noise along the genome and the predicted
reduction map $B(x)$ does not fit the data well. However, we note that under
very strong selection ($s=0.05$), $R^2$ is reduced. This is like due to very
strong selection having less localized effects and impacting overall
genome-wide diversity \parencite{Santiago1995-hx,Robertson1961-ho}.

Second, both our implementation of classic BGS theory and our B' method
accurately infer the average selection coefficient under strong selection
(Figure \ref{fig:figure-2}, middle row). However, when selection was weak, the
classic BGS model erroneously estimates strong selection and a very low
mutation rate. By contrast, our B' method estimated selection coefficients much
closer to their true value. A minor discrepancy occurs around $2Ns = 1$, likely
due to the sensitivity of mutations in this region to selective interference
(discussed in Section \ref{sec:chrom-sims}). Because we only simulated fixed
selection coefficients and five chromosomes, we only assessed the accuracy of
average selection coefficients and not the DFE probabilities.

Third, we find slight biases in mutation rated estimated by both our B' and the
classic BGS methods (Figure \ref{fig:figure-2}, bottom row). However, mutation
rate estimates based on our B' method are more accurate than classic BGS theory
across a range of selection coefficients. Overall, the bias in estimated
mutation rates suggests that benchmarking genome-wide negative selection models
based on their agreement with pedigree-based rate estimates may not be
appropriate. When BGS is not functioning, either due to weak selection or a low
rate of deleterious mutations (Figure \ref{fig:figure-2}, right column), all
estimates deteriorate. This is understandable, as the overall signal from
linked selection weakens relative to drift-based noise. We should note, though,
that this is an unlikely region of parameter space and this issue can be
readily diagnosed from the low $R^2$ values.

Finally, we tested the sensitivity of our method to violations of two
assumptions---constant population size and recessivity of selective
effects---through simulations. Using an expansion rate and timing as estimated
for Europeans by \textcite{Gutenkunst2009-pg}, we found that a population
expansion had minimal impact on parameter estimates (Supplementary Materials
Section \ref{supp:sim-assum}). We did not test the influence of population
bottlenecks, since parameter estimates of out-of-Africa bottlenecked
populations (CHB and CEU) did not differ much from YRI estimates (see below).
However, these simulations did find that selection coefficient estimates were
downwardly biased when deleterious mutations were recessive and weak.
Recessivity did not bias selection coefficient estimates when selection was
strong; this is expected since strongly deleterious mutations are predominantly
are sufficiently rare that they are not often found in a homozygote. Overall,
the sensitivity to recessivity is likely due to the ratchet rate being
overestimated when selection is weak and recessive mutations drift up to higher
frequency and experience stronger than expected selection in homozygotes.

@@

Our method estimates the distribution of fitness effects (DFE) for each feature
class. While CADD-based models only have a single conserved feature class (e.g.
CADD 6\%), feature-based models can have multiple feature classes under varying
levels of selective constrain. However, overlapping features (e.g. a basepair
that is annotated as both PhastCons and coding sequence) must be assigned to
one category or the other. Since this assignment impacts DFE estimates, we fit
both of the two alternative models. First, a \emph{PhastCons Priority} model,
where genic features that overlap PhastCons regions are classified as
PhastCons, and all remaining coding basepairs are labeled as CDS. Second, a
\emph{Feature Priority} model, where all coding basepairs are assigned to CDS,
and the PhastCons class catches the remaining highly-conserved non-genic
regions.

@@

Our PhastCons Priority model fits marginally better than the CADD 6\% model
consistently across all three populations (Supplementary Materials Table
\ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
more DFE parameters, and thus can fit finer-scale differences among the four
feature types. By contrast, the CADD 6\% track must fit patterns of diversity
to conserved sites under a single DFE. The CADD 8\% model fit slightly worse
than the CADD 6\% across all populations, and the Feature Priority model had
the poorest-fit.

@@

By contrast, full annotation models fit poorly under classic BGS theory, and
lead to unreasonable parameter estimates (discussed in Section XXX).
Additionally, when sparse annotation models contain genomic features that are
likely under weak constraint (such as introns and UTRs), models fit worse under
classic BGS theory than our B' method (Figure \ref{fig:figure-2}A). However,
among the CADD annotation models, the goodness-of-fit is nearly identical
between B' and classic BGS methods. This behavior is what we would expect given
that the CADD models contain only the most pathogenic sites, which are \emph{a
priori} very likely under the strong selection domain under which B' and
classic BGS theory agree.

>>>>

    Across the human genome, there are large-scale fluctuations in genetic
    diversity caused by the indirect effects of selection. This can be thought
    of as a ``linked selection signal" that reflects the impact of selection
    varying according to the placement of functional regions and recombination
    rates along the genome. Previous work has shown that negative selection
    against the steady influx of new deleterious mutations into conserved
    regions is the predominant mode of selection in humans. However, the
    theoretic model that underpins these results, classic Background Selection
    theory, is only applicable when new mutations are so deleterious that they
    cannot fix in the population. Here, we develop a statistical method based
    on a quantitative genetics view of the linked selection, which models the
    effects of weak draft created according to how polygenic additive fitness
    variance is distributed along the genome. We use a recent model that
    jointly predicts the equilibrium fitness variance and substitution rates
    due to both strong and weakly deleterious mutations, we estimate the
    distribution of fitness effects (DFE) and mutation rate across three human
    populations. While our model can accommodate weaker selection, we initially
    find evidence across three human populations of very strong selection
    against deleterious mutations consistent with previous work. However, the
    corollary predicted substitution rates for conserved regions are
    unreasonably low, and in disagreement with observed rates. We hypothesize
    this could be due to selected sites experiencing a further diminished
    population size due to selective interference. When we account for this in
    our method, we find evidence of weakly deleterious mutations in conserved
    regions which brings the predicted substitution rate into agreement with
    observations. However, these models lead to implausibly large mutation rate
    estimates. Overall, while our model of the genomic linked selection signal
    brings us a step towards uniting population and quantitative genetic
    selection models with the substitution process, our work suggests
    considerable uncertainty remains about the processes generating fitness
    variance in humans.

@@

% Similar to previous work, we aim to characterize selection processes through
% the signal left on nucleotide diversity at linked neutral sites. Although
% nucleotide diversity is a less informative statistic than the full site
% frequency spectrum, it is much easier to derive theoretic models of how linked
% selection alters expected pairwise coalescence times than to describe the
% entire genealogy. Still, inferences drawn from the linked selection signal
% about selective processes are only as accurate as the underlying theoretic
% model. Since we base our method on a fundamentally different theoretic approach
% to modeling linked selection, we briefly introduce the theory was extend to
% apply across whole-genomes.

Our work extends quantitative genetic models of linked selection
\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
which model the effects linked selection in terms of polygenic additive fitness
variance ($V_A$). These linked selection models stem from
\textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
fitness variation causes weak genetic draft. At the individual level Robertson
considered, selection generates autocorrelation as offspring from large
families tend to beget many descendents themselves (and likewise with small
families) when fitness is heritable. This same autocorrelation occurs at the
genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as the
perturbations to a neutral allele's trajectory from its particular fitness
background tend to occur in the same direction across generations until the
background recombines off. These models quantify the total impact of the
autocorrelation generated by weak draft in terms of what we think of as a
\emph{draft-effective} population size $N_d$. The key insight is that in the
long-run, steady, weak draft created by the presence of additive genetic
fitness variance $V_A > 0$ acts like an extra variance in offspring number
would under pure drift \parencite{Wright1938-tv}. However, due to
autocorrelation, the cumulative effect of this fitness variance is inflated by
a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the expected
total variance in reproductive success a neutral mutation experiences over its
lifetime in a system with weak selection draft.

@@

\begin{align}
    \label{eq:polygenic_Ne}
    N_d \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
\end{align}
%
where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Materials Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. The autocorrelation function for a neutral allele associated
with segment $i$ is $C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the
recombination fraction to the segment and $\kappa_i$ is the rate that the
associated fitness variance decays due to selective dynamics. Then, the
cumulative autocorrelation over the lifespan of the allele is,

\begin{align}
    \label{eq:Q}
    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \nonumber \\
        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)}.
\end{align}
%
(see Supplementary Materials Equation \ref{supp-eq:Qinf}). This general equation can accommodate
models of polygenic selection as long as the equilibrium additive fitness
variation $V_{A,i}$ can be specified and the change in variance due to
selection can be approximated as a geometric decay, i.e. $\Delta V_{A,i} =
-\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This
is usually a reasonable assumption since within-generation selection removes a
fraction of phenotypic variation from the population, and some fraction of that
is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.

@@

\begin{align}
    N_d = N \exp \left( - \sum_i^S \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
\end{align}
%
which is identical to the genome-wide model of background selection used in
previous studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}.
Thus, the classic background selection model is a special case of the more
general theory of Santiago and Caballero (\citeyear{Santiago2016-mu}), which
they had shown previously (\citeyear{Santiago1998-bs}).

However, when new mutations are only weakly deleterious, they can drift up in
frequency before their eventual loss or fixation. At this point, the number of
deleterious mutations per haplotype is no longer well-approximated by the
deterministic mutation-selection-balance theory; their dynamics are strongly
influenced by stochastic perturbations due to both drift and draft. In this
weak selection regime, classic BGS no longer accurately predicts levels of
linked diversity
\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
Moreover, selection against weakly deleterious mutations alters the
genealogies, such that they are no longer well-approximated by a rescaled
neutral coalescent as assumed under the classic BGS model
\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. To further
complicate matters, the distribution of the number of deleterious mutations
(and its corresponding fitness distribution) is no longer a stationary Poisson
distribution, instead becoming a traveling wave
\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
population fitness (since we ignore beneficial mutations). When this occurs,
individuals in the least-loaded class are stochastically lost, fixing those
mutations with a click of ``Muller's ratchet"
\parencite{Muller1964-ki,Charlesworth1997-qn}. Unfortunately, determining the
rate of Muller's ratchet is another difficult, open problem
\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz,Higgs1995-xc} related to
Hill--Robertson interference \parencite{Felsenstein1974-xm}.

Under qualitative genetic models of linked selection, these complications can
be avoided by finding a general expression for the equilibrium $V_A$ applicable
for both strong and weakly deleterious mutations. \textcite{Santiago2016-mu}
suggest that equilibrium fitness variance is lower than predicted by
$V_{a}^{BGS}$ once weakly deleterious mutations begin to have an appreciable
rate of fixation, $R > 0$, since each substitution (or click of Muller's
``ratchet") removes a segregating site and its heterozygosity becomes zero.
Thus, the steady-state genic variance under mutation and negative selection is
(omitting the segment index $i$ for clarity),

@@

where the condition $V_a \ge 0$ is met when the probability of fixation is less
than or equal to the neutral fixation probability of $\nicefrac{1}{2N}$, which
is held when considering deleterious mutations. This equation connects the
equilibrium additive genic fitness variance to the flux of new variation in
from new deleterious mutations, and the flux out due to their substitution (and
the decline in mean population fitness). When $R=0$, selection is so strong it
cannot fix, and the equilibrium fitness variation is due entirely to young rare
mutations before their extinction $V_a = V_a^{BGS} \approx Us$. Santiago and
Caballero derive Equation \eqref{eq:Va} through Fisher's Fundamental Theorem of
Natural Selection, but we find an alternative proof (\cite{Higgs1995-xc}; see
Supplementary Materials Section \ref{supp:weak-strong}). We also find that the
steady-state additive genic variance in Equation \eqref{eq:Va} results from
diffusion models with a flux of mutations into discrete sites
(\cite{Kimura1969-jw}).

@@

Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
ratchet rate for a single segment under general mutation-selection processes,
we next extended their model so that it could be applied to whole-genome data.
Our software method \texttt{bgspy} numerically solves Equations
\eqref{eq:main_eqns} to compute the equilibrium additive genic fitness variance
($\widetilde{V}_a$) and ratchet rate ($\widetilde{R}$) across grids of mutation
rates and selection coefficients. This is done for each pre-specified segment
in the genome that may be under negative selection (e.g. coding sequences or
UTRs). We call the set of predicted reductions across these grids the B' maps
(to distinguish them from McVicker's B maps); these can be used then to find
the equilibrium reduction factor $B(x)$ for any genomic position $x$.

We validated our predicted B' reduction maps with realistic chromosome-scale
forward simulations of negative selection in conserved regions. We find that
our B' maps and the classic BGS theory B maps closely match simulations when
selection is strong (top row of Figure \ref{fig:figure-1}E), apart from slight
discrepancies in low recombination regions (Figure \ref{fig:figure-1}E).
Second, we find our theory is vastly more accurate than the classic BGS when
selection is very weak ($2N_e s \ll 1$; bottom row of Figure
\ref{fig:figure-1}E). Across all mutation and selection parameters simulated,
the relative error of the classic B maps is 14.6\% whereas the relative error
in the new B' maps is 5\%. Nearly all of this error is in the drift-barrier
domain ($2N_e s \approx 1$ domain); for strong and weak selection, the mean
squared error between simulations and B' maps is close to the theoretic lower
bound of $\V(B) \approx \nicefrac{2}{9n}$ set by the coalescence variance
\parencite{Tajima1983-gu}.

We hypothesized this error in the drift-barrier domain may be due to selective
interference between segments that is not taken into account when we
numerically solve Equations \eqref{eq:main_eqns} independently for each
segment. In particular, we use a fixed drift-effective $N=1,000$ corresponding
to the number of diploids in the simulations rather than $B(x) N$, which is an
approximation to the draft-effective population size experienced by a selected
segment at position $x$. To test this, we implemented a ``locally-rescaled"
version of the B' maps, which uses $B(x)N$ as the population size. We find the
locally-rescaled B' maps reduce the relative error from 5\% to 0.4\% and mean
squared error (Figure \ref{fig:figure-1}D, dashed colored lines), but does not
entirely eliminate the error in the $2Ns \approx 1$ domain.

Finally, we hypothesized that the remaining error could be due to the build up
of negative linkage disequilibria between selected sites due to Hill--Robertson
interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. By
calculating the sum of all linkage disequilibria pairs in our chromosome-wide
simulations, we find such a build up of negative LD around $2Ns \approx 1$
(Figure \ref{fig:figure-1}D, top row). As $Ns \to 0$, the variance in LD
inflates as expected \parencite{Ohta1969-ae,Hill1968-ue}. Overall, the build up
of negative LD is consistent our view that the equilibrium fitness variance
modeled by the SC16 theory is the additive \emph{genic} fitness variance, which
differs from the additive genetic variance by the sum of linkage disequilibria
between selected sites (i.e. $\delta_{LD} = s^2 \sum_{i\ne j} D_{i,j}$ where
$D_{i,j}$ is the LD between sites $i$ and $j$). However, according to theory,
the reductions in diversity should be determined by levels of additive genetic
fitness variance that include the contribution of LD (Supplementary Materials
Section \ref{supp:theory}).

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=0.8\textwidth]{figures/figure_2.pdf}

    \caption{Comparison of parameter estimates using classic BGS theory (green
        lines) with our new B' method (blue lines) across both full and sparse
        track types (dark versus light hue), and different mutation rates
        (columns). In the first row, $R^2$ between predictions and observations
        increases with selection intensity. Both classic BGS and B' methods fit
        the data equally well with sparse tracks, though the estimated
        selection coefficients are inaccurate when selection is weak (second
        row). Mutation rate estimates (third row) are more accurately estimated
        by the B' method than classic BGS, but overall show slight biases.
        Overall, classic BGS methods break down as expected when full-coverage tracks
    are used, since it cannot accommodate weak selection and neutrality. }

  \label{fig:figure-2}
\end{figure}

@@

First, we find the $R^2$ between predicted and simulated megabase-scale
diversity serves as a measure of the strength of the linked selection signal in
genome-wide data. The $R^2$ increases with the intensity of selection against
new deleterious mutations and mutation rate (Figure \ref{fig:figure-2}, top
row). Under just drift or weak negative selection, the variance in diversity is
driven by unstructured coalescence noise along the genome and the predicted
reduction map $B(x)$ does not fit the data well. Under very strong selection
($s=0.05$), $R^2$ is reduced; this is likely due to very strong selection
having less localized effects and impacting overall genome-wide diversity
\parencite{Santiago1995-hx,Robertson1961-ho}.

Second, both our implementation of classic BGS theory and our B' method
accurately infer the average selection coefficient under strong selection
(Figure \ref{fig:figure-2}, middle row). However, when selection was weak, the
classic BGS model erroneously estimates strong selection and a very low
mutation rate. By contrast, our B' method estimated selection coefficients much
closer to their true value. A minor discrepancy occurs around $2Ns = 1$, likely
due to the sensitivity of mutations in this region to selective interference
(these do not use local-rescaling). Because we only simulated fixed selection
coefficients and five chromosomes, we only assessed the accuracy of average
selection coefficients and not the DFE probabilities.

Third, we find slight biases in mutation rated estimated by both our B' and the
classic BGS methods (Figure \ref{fig:figure-2}, bottom row). However, mutation
rate estimates based on our B' method are more accurate than classic BGS theory
across a range of selection coefficients. Overall, the bias in estimated
mutation rates suggests that benchmarking genome-wide negative selection models
based on their agreement with pedigree-based rate estimates may not be
appropriate. When BGS is not occurring, either due to weak selection or a low
rate of deleterious mutations (Figure \ref{fig:figure-2}, right column), all
estimates deteriorate. This is understandable, as the overall signal from
linked selection weakens relative to drift-based noise. We should note, though,
that this is an unlikely region of parameter space and this issue can be
readily diagnosed from the low $R^2$ values. Finally, we find in additional
tests that our estimates are robust to demographic expansions but sensitive
when mutuations have recessive effects, since our model assumes additive
effects (see Supplementary Materials Section \ref{supp:sim-assum}). We did not
test the influence of population bottlenecks, since parameter estimates of
out-of-Africa bottlenecked populations (CHB and CEU) did not differ much from
YRI estimates (see below).

@@

Our method estimates a distribution of fitness effects (DFE) for each feature
class. While CADD-based models only have a single conserved feature class (e.g.
CADD 6\%), feature-based models can have multiple feature classes under varying
levels of selective constrain. However, overlapping features (e.g. a basepair
that is annotated as both PhastCons and coding sequence) must be assigned to
one category or the other. Since this assignment impacts DFE estimates, we fit
both of the two alternative models. First, a \emph{PhastCons Priority} model,
where genic features that overlap PhastCons regions are classified as
PhastCons, and all remaining coding basepairs are labeled as CDS. Second, a
\emph{Feature Priority} model, where all coding basepairs are assigned to CDS,
and the PhastCons class catches the remaining highly-conserved non-genic
regions.

@@

% Our PhastCons Priority model fits marginally better than the CADD 6\% model
% consistently across all three populations (Supplementary Materials Table
% \ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
% more DFE parameters, and thus can fit finer-scale differences among the four
% feature types. By contrast, the CADD 6\% track must fit patterns of diversity
% to conserved sites under a single DFE. The CADD 8\% model fit slightly worse
% than the CADD 6\% across all populations, and the Feature Priority model had
% the poorest-fit.

@@

By contrast, full annotation models fit poorly under classic BGS theory, and
lead to unreasonable parameter estimates. Additionally, when sparse annotation
models contain genomic features that are likely under weak constraint (such as
introns and UTRs), models fit worse under classic BGS theory than our B' method
(Figure \ref{fig:figure-2}A). However, among the CADD annotation models, the
goodness-of-fit is nearly identical between B' and classic BGS methods. This
behavior is what we would expect given that the CADD models contain only the
most pathogenic sites, which are \emph{a priori} very likely under the strong
selection domain under which B' and classic BGS theory agree.
